---
title: "Cleaning Data Week 4 Assignment"
author: "LloydW"
date: "25/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Read in the test and train data files.

```{r}
X_test <- read.table("H:/R/Training/UCI HAR Dataset/test/X_test.txt")
Y_test <- read.table("H:/R/Training/UCI HAR Dataset/test/y_test.txt")
Body_acc_x_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/body_acc_x_test.txt")
Body_acc_y_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/body_acc_y_test.txt")
Body_acc_z_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/body_acc_z_test.txt")
Body_gyro_x_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/body_gyro_x_test.txt")
Body_gyro_y_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/body_gyro_y_test.txt")
Body_gyro_z_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/body_gyro_z_test.txt")
Total_acc_x_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/total_acc_x_test.txt")
Total_acc_y_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/total_acc_y_test.txt")
Total_acc_z_test <- read.table("H:/R/Training/UCI HAR Dataset/test/Inertial Signals/total_acc_z_test.txt")
X_train <- read.table("H:/R/Training/UCI HAR Dataset/train/X_train.txt")
Y_train <- read.table("H:/R/Training/UCI HAR Dataset/train/y_train.txt")
Body_acc_x_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/body_acc_x_train.txt")
Body_acc_y_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/body_acc_y_train.txt")
Body_acc_z_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/body_acc_z_train.txt")
Body_gyro_x_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/body_gyro_x_train.txt")
Body_gyro_y_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/body_gyro_y_train.txt")
Body_gyro_z_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/body_gyro_z_train.txt")
Total_acc_x_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/total_acc_x_train.txt")
Total_acc_y_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/total_acc_y_train.txt")
Total_acc_z_train <- read.table("H:/R/Training/UCI HAR Dataset/train/Inertial Signals/total_acc_z_train.txt")
```

Read activity labels table

```{r}
ActivityLabels <- read.table("H:/R/Training/UCI HAR Dataset/activity_labels.txt")
ActivityLabels <- ActivityLabels %>% rename(Activity_Code = V1, Activity_Name = V2)
ActivityLabels
```

Read features table and give sensible column names

```{r}
Features <- read.table("H:/R/Training/UCI HAR Dataset/features.txt")
Features <- Features %>% rename(Feature_Code = V1, Feature_Names = V2) %>% mutate(Feature = make.names(Feature_Names)) %>% select(Feature_Code, Feature)
Features
```

Read subject data tables

```{r}
Subject_test <- read.table("H:/R/Training/UCI HAR Dataset/test/subject_test.txt")
Subject_train <- read.table("H:/R/Training/UCI HAR Dataset/train/subject_train.txt")
```

Rename columns in Subject_test and Y_test tables

```{r}
Subject_test <- Subject_test %>% rename(Subject = V1)
Y_test <- Y_test %>% rename(Activity = V1)
```

Rename V1, V2, ... in X_test with feature names

```{r}
colnames(X_test) <- Features$Feature
X_test
```

Bind subjects, activities and measurements

```{r}
Test_data <- cbind(Subject_test, Y_test, X_test) 
Test_data
```

Makes column names unique and keep only columns that are means or standard deviations

```{r}
colnames(Test_data) <- make.names(colnames(Test_data), unique = TRUE)
Test_data_final <- Test_data %>% select(Subject, Activity, matches("mean|std"), -contains("Freq"))
Test_data_final
```

Repeat same process for the train data


```{r}
Subject_train <- Subject_train %>% rename(Subject = V1)
Y_train <- Y_train %>% rename(Activity = V1)
colnames(X_train) <- Features$Feature
Train_data <- cbind(Subject_train_new, Y_train, X_train)
colnames(Train_data) <- make.names(colnames(Train_data), unique = TRUE)
Train_data_final <- Train_data %>% select(Subject, Activity, matches("mean|std"), -contains("Freq"))
Train_data_final
```

Combine test and train tables

```{r}
Data_final <- rbind(Test_data_final, Train_data_final)
```

Calculate means for each person and activity

```{r}
Summarised_by_mean <- Data_final %>% group_by(Subject, Activity) %>% summarise_all(funs(mean))
Summarised_by_mean
```

Export as text file

```{r}
Summarised_by_mean %>% write.table("Summarised_by_mean.txt", row.name=FALSE)
```

```{r}
knitr::purl("Cleaning data week 4 assignment.Rmd", "Run_Analysis.R", documentation = 2)
```




